---
title: "Sustainability and Social Equity Data Exploration"
author: "Sarina Singh Khaira"
date: "03/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

```{r include = FALSE}
library(here)
library(tidyverse)
library(RColorBrewer)
library(countrycode)
library(snakecase)
library(extrafont)
```


```{r include = FALSE}
bump_theme <- function() {

  # Colors
  color.background = "white"
  color.text = "grey40"

  # Begin construction of chart
  theme_bw(base_size=15) +

    # Format background colors
    theme(panel.background = element_rect(fill=color.background,
                                          color=color.background)) +
    theme(plot.background  = element_rect(fill=color.background,
                                          color=color.background)) +
    theme(panel.border     = element_rect(color=color.background)) +
    theme(strip.background = element_rect(fill=color.background,
                                          color=color.background)) +

    # Format the grid
    theme(panel.grid.major.y = element_blank()) +
    theme(panel.grid.minor.y = element_blank()) +
    theme(axis.ticks       = element_blank()) +

    # Format the legend
    theme(legend.position = "none") +

    # Format title and axis labels
    theme(plot.title       = element_text(color=color.text, size=15, face = "bold")) +
    theme(axis.title.x     = element_blank()) +
    theme(axis.title.y     = element_blank()) +
    theme(axis.text.x      = element_text(size=10, vjust=0.5, hjust=0.5,
                                          color = color.text, face = "bold")) +
    theme(axis.text.y      = element_blank()) +
    theme(strip.text       = element_text(face = "bold")) +

    # Plot margins
    theme(plot.margin = unit(c(0.35, 0.2, 0.3, 0.35), "cm"))
}
```


```{r include = FALSE}
gci_rankings <- read_csv(here("clean_data/gci_rank_decade.csv"))
```
For the purposes of this exploration I will include the 10 top ranking countries for GCI.

```{r include = FALSE}
# create vector of countries of interest
gci_top_10_countries <- gci_rankings %>%
  filter(year == 2018 & gci_rank <= 10) %>%
  arrange(gci_rank) %>%
  pull(country)

# filter countries of interest for rank over time
gci_rankings_top_10 <- gci_rankings %>%
  filter(country %in% gci_top_10_countries) %>%
  mutate(country = fct_relevel(country, gci_top_10_countries),
         year = as.factor(year)) %>%
  select(country, year, gci_rank)

```

```{r}
library(ggbump)
library(ggflags)
data <- gci_rankings_top_10 %>%
  # mutate(country = countrycode(country, origin = "country.name", destination = "iso2c")) %>%
    filter(year %in% c(2010:2018)) 

data %>% 
  ggplot(aes(year, gci_rank, group = country, colour = country, fill = country)) +
  geom_bump(aes(smooth = 20), size = 1.5, lineend = "round", alpha = 0.4) +
  geom_bump(data = data %>% filter(country == "United Kingdom"), 
            colour = "#d81159", size = 1.5, lineend = "round", linejoin = "bevel") +
#  geom_point(alpha = 0.4, size = 4) +
  scale_y_reverse(breaks = 1:10) +
# scale_colour_manual(values = mycolors) +
  scale_color_viridis_d(direction = -1, option = "plasma") +
  theme_minimal() +
  bump_theme() +
# add labels for 2010 rankings  
  geom_text(data = data %>% filter(year == 2010), 
            aes(label = paste(gci_rank, country), x = 0.5), 
            hjust = 1, colour = "grey40", fontface = "bold", size = 3) +
# add labels for 2018 rankings  
  geom_text(data = data %>% filter(year == 2018), 
            aes(label = paste(gci_rank, country), x = 9.5), 
            hjust = 0, colour = "grey40", fontface = "bold", size = 3) +
  scale_x_discrete(breaks = 2010:2018, expand = c(.28, .45)) +
  ggtitle("Country Rankings of Global Competitiveness Index") +
  ylim(c(12, 0))
```




```{r echo = FALSE}
gci_rankings_top_10 %>%
  filter(year %in% c(2010:2018)) %>%
  ggplot() +
  aes(x = year, y = gci_rank, group = country, colour = country) +
# plot lines for all countries
  geom_line(alpha = 0.3, size = 1, lineend = "round", linejoin = "bevel") +
# plot UK line to highlight
  geom_line(data = gci_rankings_top_10 %>% filter(country == "United Kingdom" & year %in% c(2010:2018)), 
            aes(x = year, y = gci_rank), 
            colour = "#d81159", size = 1.5, lineend = "round", linejoin = "bevel") +
#  geom_point(alpha = 0.4, size = 4) +
  scale_y_reverse(breaks = 1:10) +
# scale_colour_manual(values = mycolors) +
  scale_color_viridis_d(direction = -1, option = "plasma") +
  theme_minimal() +
  bump_theme() +
# add labels for 2010 rankings  
  geom_text(data = gci_rankings_top_10 %>% filter(year == 2010), 
            aes(label = paste(gci_rank, country), x = 0.5), 
            hjust = 1, colour = "grey40", fontface = "bold", size = 3) +
# add labels for 2018 rankings  
  geom_text(data = gci_rankings_top_10 %>% filter(year == 2018), 
            aes(label = paste(gci_rank, country), x = 9.5), 
            hjust = 0, colour = "grey40", fontface = "bold", size = 3) +
  scale_x_discrete(breaks = 2010:2018, expand = c(.28, .45)) +
  ggtitle("Country Rankings of Global Competitiveness Index") +
  ylim(c(12, 0))

```


# Heatmap of top countries and each pillar ranking

```{r}
gci_soc_env <- read_csv(here("clean_data/gdp_gci_clustering.csv"))
```

```{r}
# obtain the gci rankings since 2007 in order of gci
gci_top_10_countries <- gci_rankings %>%
  filter(year == 2018 & gci_rank <= 10) %>%
  arrange(gci_rank) %>%
  pull(country_code)

gci_top_10_country_names <- countrycode(gci_top_10_countries, origin = "iso3c", destination = "country.name")

# select out index values for 12 pillars
pillars <- gci_soc_env %>%
  # pillar values 
  select(country_code, starts_with("p_"), gci_overall) %>%
  # filter out countries of interest
  filter(country_code %in% gci_top_10_countries) %>%
  # add full country names
  mutate(country_name = countrycode(country_code, 
                                    origin = "iso3c", 
                                    destination = "country.name")) %>%
  # remove gci index total
  select(-gci_overall) %>%
  # pivot so that all pillar scores are in one column
  pivot_longer(cols = -c(country_name, country_code),
               names_to = "pillar",
               values_to = "score") %>%
  # modify pillar labels so that they are understandable for plot
  mutate(pillar_labs = (str_extract(pillar, "[a-z]*$")),
         pillar_labs = to_sentence_case(pillar_labs),
         pillar_labs = recode(pillar_labs,
           "Ict" = "ICT adoption",
           "Macroeconomic" = "Macroeconomic\nstability", 
           "Product" = "Product market", 
           "Labour" = "Labour market", 
           "Financial" = "Financial system", 
           "Market" = "Market size", 
           "Business" = "Business\ndynamism", 
           "Innovation" = "Innovation", 
           .default = pillar_labs
         ))


```



```{r gci_heatmap}
# function to replace the pillar names so that they are sequential for plot
for (x in 1:9){
  
  string <- paste("_", x, "_", sep = "")
  
  string_replace <- paste("_0", x, "_", sep = "")
  
  pillars <-  pillars %>%
  mutate(pillar = str_replace_all(pillar, string, string_replace))
  
}

# ORder the pillars in their order for factor reordering
pillar_order <- pillars %>%
  arrange(pillars) %>%
  distinct(pillar_labs) %>%
  pull(pillar_labs)

# Vector of colours of the 4 major categories of pillar
colours <- c(rep("#e04059", 4), rep("#eb701c", 2), rep("#4f529c", 4), rep("#38a380", 2))

# Create a tibble for secondary x axis labels of major categories
# second_x_axis = tibble(
#   x_posi = c(0.5, 4.5, 6.5, 10.5),
#   pillar_cat = c("Enabling\nEnvironment", "Human\nCapital", "\n Markets", "Innovation\nEcosystem")
# )


# Heatmap og 12 pillars per 12 countries

heatmap <- 
pillars %>%
  #relevel pillars so they are plotted in order
  mutate(pillar_labs = fct_relevel(pillar_labs, pillar_order),
         country_name = fct_relevel(country_code, gci_top_10_countries)
        ) %>%
  ggplot() +
  geom_tile(aes(y = pillar_labs, 
                x = country_name, 
                fill = score),
            colour = "white", 
            size = 0.5) +
  # make the tiles square
  coord_equal() +
  # set the colour scale
  scale_fill_distiller(palette = "Purples",  
                       name = "Score", 
                       direction = 1,
                       guide = guide_colourbar(barwidth = 0.9, 
                                               barheight = 2.5,
                                               ticks = F,
                                               breaks = c(60, 100),
                                               labels = c(60, 100),
                                               label.position = "left",
                                               #title.position = "bottom",
                                               title.hjust = 1,
                                               title.vjust = 1.5)) +
# #  geom_text(data = second_x_axis, 
#             aes(x = x_posi, y = 10, label = pillar_cat), 
#             vjust = -0.8,
#             size = 2.5,
#             colour = colours[c(1, 5, 7, 11)],
#             fontface = "bold",
#             hjust = 0) +
#     theme(plot.margin = unit(c(1, 0.2, 0.3, 0.35), "cm")) +
#  expand_limits(y = 13, x = 13) +
  # add lines that group the pillars together
  # enabling environment
  annotate(
    geom = "line", y = c(0.6, 4.4), x = 0.3, col = colours[1], size = 2.5
  ) +
  # human capital
  annotate(
    geom = "line", y = c(4.6, 6.4), x = 0.3, col = colours[5], size = 2.5
  ) +
  # markets
  annotate(
    geom = "line", y = c(6.6, 10.4), x = 0.3, col = colours[8], size = 2.5
  ) +
  # innovation ecosystem
  annotate(
    geom = "line", y = c(10.6, 12.4), x = 0.3, col = colours[11], size = 2.5
  ) +
  # styling to remove all gridlines and axis
  theme_minimal() +
  # x axis label rotate 
  theme(axis.text.x = element_text(angle = 45, 
                                   #hjust = 1, # left justification
                                   #vjust = 0.5, 
                                   face = "bold"),
        
        axis.text.y = element_text(face = "bold",
                                   colour = colours),
        axis.title = element_text(size = 10,
                                  face = "bold"),
        axis.ticks = element_blank(),
        
        legend.title = element_text(size = 10, 
                                    face = "bold", 
                                    color = "grey40"),
        
        legend.text = element_text(face = "bold", 
                                   color = "grey40")) +
#  scale_x_discrete(position = "top") +
    labs(x = "Country Code", 
       y = NULL)

ggsave(here('plot_images/gci_heatmap.png'), heatmap, height = 4)
heatmap
```

## Plot of GCI Rankings and GCSI rankings/scores

```{r}
gci_soc_env %>%
  slice_max(gci_overall, n = 30) %>%
  mutate(index_diff = gci_overall - gsci_score,
    country_code = fct_reorder(country_code, index_diff)) %>%
  select(country_code, gci_overall, gsci_score) %>%
  pivot_longer(cols = -country_code,
               names_to = "index", 
               values_to = "score") %>%
  ggplot() +
  aes(x = score, y = country_code) +
  geom_line(aes(group = country_code)) +
  geom_point(aes(color = index), size = 4) +
  theme(legend.position = "none") +
  theme_light() 
```

```{r}
gsci_rankings <- read_csv(here("clean_data/gsci_all.csv")) %>%
  select(gsci_rank, gsci_score, country_code) %>%
  rename(rank = gsci_rank,
         score = gsci_score) %>%
  mutate(index = "gsci")

gci_gsci_rankings <- gci_rankings %>%
  filter(year == "2018") %>%
  select(gci_rank, gci_value, country_code) %>%
    rename(rank = gci_rank,
         score = gci_value) %>%
  mutate(index = "gci") %>%
  bind_rows(gsci_rankings)


```


```{r}

ranking_countries <- gci_gsci_rankings %>%
  filter(rank <= 10) %>%
  distinct(country_code) %>%
  pull()

gci_gsci_rankings %>%
  filter(country_code %in% ranking_countries) %>%
  group_by(index) %>%
  mutate(ordering = row_number(rank)) %>%
  ggplot() +
  aes(x = index, y = (ordering), group = country_code) +
  geom_line() +
  geom_label(aes(label = paste0(rank, " ", country_code))) +
  scale_y_reverse() +
  theme_void() +
  theme(legend.position = "none") +
  scale_colour_viridis_d(option = "plasma") +
  annotate()
  
```

```{r}
library(ggbump)
gci_gsci_rankings_top_20 <- gci_gsci_rankings %>%
  filter(country_code %in% ranking_countries) %>%
  group_by(index) %>%
  mutate(ordering = row_number(rank),
         country_code = countrycode(country_code, "iso3c", "country.name")) 

gci_gsci_rankings_top_20 %>%
  ggplot() +
  aes(x = index, y = rank, group = country_code) +
# plot lines for all countries
  geom_line(alpha = 0.3, size = 1, lineend = "round", linejoin = "bevel") +
# plot UK line to highlight
#  geom_line(data = gci_rankings_top_10 %>% filter(country == "United Kingdom" & year %in% c(2010:2018)), 
#            aes(x = year, y = gci_rank), #
#            colour = "#d81159", size = 1.5, lineend = "round", linejoin = "bevel") +
#  geom_point(alpha = 0.4, size = 4) +
  scale_y_reverse() +
# scale_colour_manual(values = mycolors) +
  scale_color_viridis_d(direction = -1, option = "plasma") +
  theme_minimal() +
  bump_theme() +
  theme(panel.grid = element_blank()) +
# add labels for 2010 rankings  
  geom_text(data = gci_gsci_rankings_top_20 %>% filter(index == "gci"), 
            aes(label = paste(rank, country_code), x = 0.99, y = rank), 
            hjust = 1, colour = "grey40", fontface = "bold", size = 3) +
# add labels for 2018 rankings  
  geom_text(data = gci_gsci_rankings_top_20 %>% filter(index == "gsci"), 
            aes(label = paste(rank, country_code), x = 2.01), 
            hjust = 0, colour = "grey40", fontface = "bold", size = 3) +
#    scale_x_discrete(breaks = 2010:2018, expand = c(.28, .45)) +
#  ggtitle("Country Rankings of Global Competitiveness Index") +
  ylim(c(15, 0))
```


```{r}
gci_gcsci_plot <- gci_gsci_rankings_top_20 %>%
  select(country_code, index, rank) %>%
  unique() %>%
  pivot_wider(names_from = "index",
              values_from = "rank",
              values_fn = max) %>%
  ggplot() +
  geom_sigmoid(
    aes(x = -0.5, xend = 0.5, y = gci, yend = gsci, group = country_code), size = 1.5, alpha = 0.8, color = "grey40") +
  geom_sigmoid(aes(x = -0.5, xend = 0.5, y = 8, yend = 15), colour = "#d81159", size = 1.5) +
  geom_text(aes(x = -0.5, y = gci, label = paste(gci, country_code)), hjust = 1, fontface = "bold") +
  geom_text(aes(x = 0.5, y = gsci, label = paste(gsci, country_code)), hjust = 0, fontface = "bold") +
  annotate(geom = "text", x = 0.5, y = 0, label = "GSCI", hjust = 0, size = 5, fontface = "bold") +
  annotate(geom = "text", x = -0.5, y = 0, label = "GCI", hjust = 1, size = 5, fontface = "bold") +
  scale_y_reverse() +
  scale_x_discrete(expand = c(1.5, .45)) +
  theme_void() +
   ylim(c(15, 0)) +
  theme(plot.margin = unit(c(0.35, 1, 0.3, 1), "cm"),# top right bottom left
        text = element_text(face = "bold", colour = "blue")) 

ggsave(here("plot_images/gci_gsci_rank.png"), gci_gcsci_plot, width = 8, height = 5)
```


```{r}
gci_gsci_rankings_top_20 %>%
  ggplot() +
  aes(x = index, y = rank, group = index) +
# plot lines for all countries
  geom_line(alpha = 0.3, size = 1, lineend = "round", linejoin = "bevel")
```




# Plot of GDP growth

```{r}
gdp <- read_csv(here("clean_data/gdp.csv"))

# vector of countries of interest
top_countries <- read_csv(here("clean_data/top_countries.csv"))
```

```{r}
# extract countries of interest from gdp data
gdp_top_countries <- top_countries %>%
  left_join(gdp, by = "country_code") %>%
  drop_na()
```


```{r}
# select variables of interest
gdp_tc_plot <- gdp_top_countries %>%
  select(-c(gdp_ppp, gdp_us, gdp_growth_annual, gdp_per_capita_ppp, gdp_per_capita_us)) %>%
  filter(year > 2010) %>%
  mutate(year = as.factor(year),
         country_name = countrycode(country_code, "iso3c", "country.name")) %>%
  pivot_longer(cols = starts_with("gdp"),
               names_to = "gdp_measure") 

# plot gdp growth per capita
#gdp_plot <- 
gdp_tc_plot %>%
  filter(index != "uk") %>%
  ggplot() +
  aes(x = year, 
      y = value,
      group = country_code,
      colour = index) +
  # set an abline at y = 0
  geom_abline(slope = 0, intercept = 0, col = "orange", alpha = 0.8) +
  geom_line(alpha = 0.5, size = 1.2) +
#  geom_point(aes(shape = index)) +
 facet_wrap(~country_name) +
  # add the UK without alpha
  geom_line(data = gdp_tc_plot %>% filter(index == "uk"),
            aes(x = year,
                y = value), colour = "#E04059", size = 1.2) +
  # geom_text(data = gdp_tc_plot %>% filter(index == "uk"),
  #           aes(x = 9,
  #               y = 1.2,
  #               label = country_code),
  #           colour = "#E04059") + 
  scale_colour_manual(values = c("#EB701C", "#38A380"), 
                      labels = c("Global Competitiveness Index (WEF)",
                                 "Global Sustainable Competitiveness Index"),
                      name = "Ranking Index") +
  labs(x = "Year", 
       y = "GDP per capital growth (%)", 
       title = "Annual growth in GDP per capita",
       subtitle = "Countries ranking in the top Global Competitiveness Ranking") +
   theme_bw () +
  theme(legend.position = "bottom",
        axis.text.x = element_text(angle = 45, hjust = 1), 
        # format facet header rectangle
        strip.background = element_rect(
          color="black", 
          fill="white", 
          linetype="blank"),
        # format facet header text
        strip.text.x = element_text(face = "bold"),
        # remove minor grid lines
        panel.grid.minor = element_blank()
        )

#ggsave(here("plot_images/gdp_annual_growth.png"), gdp_plot)

```

Economic growth is slowing in most countries except for Denmark. This is to be expected as high rates of economic growth are not sustainable and are typical of developing nations. The UK shows stable growth accross the last decade, though it is quite low. 


```{r}
gdp_uk <- gdp_top_countries %>%
  filter(country_code == "GBR") %>%
  pivot_longer(cols = starts_with("gdp"))

gdp_uk %>%
  ggplot() +
  aes(x = year, y = value) +
  geom_line() +
  facet_wrap(~name, scales = "free_y") +
  scale_x_continuous(breaks = c(2010:2019), labels = c(2010:2019))
```


```{r}
library(scales)

gdp_uk_plot <- gdp_uk %>%
  filter(name == "gdp_per_capita_us" |
           name == "gdp_per_capita_ppp") %>%
  ggplot() +
  aes(x = year, y = value, colour = name) +
  geom_line(size = 1.2, alpha = 0.6) +
  scale_x_continuous(breaks = c(2010:2019), labels = c(2010:2019)) +
  scale_y_continuous(labels = comma) +
  labs(x = "Year",
       y = "GDP per capita") +
  scale_colour_manual(values = c("#EB701C", "#38A380"), 
                      labels = c("Purchasing Power Parity",
                                 "US Dollar ($)"),
                      name = "GDP per capita measured in:") +
  theme_bw() +
    theme(legend.position = "bottom",
        # remove minor grid lines
        panel.grid.minor = element_blank(),
        axis.text      = element_text(size=10,
                                      color = "grey40", 
                                      face = "bold"),
        legend.title = element_text(size = 8, 
                                    face = "bold", 
                                    color = "grey40"),
        axis.title = element_text(size=8,
                                  color = "grey40", 
                                  face = "bold"))

ggsave(here("plot_images/gdp_uk.png"), gdp_uk_plot, height = 3, width = 5)
```


