---
title: "Final analysis"
author: "Sarina Singh Khaira"
date: "02/11/2021"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r include = FALSE}
library(here)
library(tidyverse)
library(RColorBrewer)
library(countrycode)
library(snakecase)
library(extrafont)
library(ggbump)
```

# How have global competitiveness rankings changed over time?

```{r include = FALSE}
# read in data for past 10 years
gci_rankings <- read_csv(here("clean_data/gci_rank_decade.csv"))

# read in competitiveness, social and environmental metrics for most recent year
gci_soc_env <- read_csv(here("clean_data/gdp_gci_clustering.csv"))

# read in global sustainable competitiveness data 
gsci_rankings <- read_csv(here("clean_data/gsci_all.csv")) %>%
  select(gsci_rank, gsci_score, country_code) %>%
  rename(rank = gsci_rank,
         score = gsci_score) %>%
  mutate(index = "gsci")

```


```{r include = FALSE}
# create custom theme for bump plot 
bump_theme <- function() {

  # Colors
  color.background = "white"
  color.text = "grey40"

  # Begin construction of chart
  theme_bw(base_size=15) +

    # Format background colors
    theme(panel.background = element_rect(fill=color.background,
                                          color=color.background)) +
    theme(plot.background  = element_rect(fill=color.background,
                                          color=color.background)) +
    theme(panel.border     = element_rect(color=color.background)) +
    theme(strip.background = element_rect(fill=color.background,
                                          color=color.background)) +

    # Format the grid
    theme(panel.grid.major.y = element_blank()) +
    theme(panel.grid.minor.y = element_blank()) +
    theme(axis.ticks       = element_blank()) +

    # Format the legend
    theme(legend.position = "none") +

    # Format title and axis labels
    theme(plot.title       = element_text(color=color.text, size=15, face = "bold")) +
    theme(axis.title.x     = element_blank()) +
    theme(axis.title.y     = element_blank()) +
    theme(axis.text.x      = element_text(size=10, vjust=0.5, hjust=0.5,
                                          color = color.text, face = "bold")) +
    theme(axis.text.y      = element_blank()) +
    theme(strip.text       = element_text(face = "bold")) +

    # Plot margins
    theme(plot.margin = unit(c(0.35, 0.2, 0.3, 0.35), "cm"))
}
```




For the purposes of this exploration I will include the 10 top ranking countries for GCI.

```{r include = FALSE}
# create vector of countries of interest
gci_top_10_countries <- gci_rankings %>%
  filter(year == 2018 & gci_rank <= 10) %>%
  arrange(gci_rank) %>%
  pull(country)

# filter countries of interest for rank over time
gci_rankings_top_10 <- gci_rankings %>%
  filter(country %in% gci_top_10_countries) %>%
  mutate(country = fct_relevel(country, gci_top_10_countries),
         year = as.factor(year)) %>%
  select(country, year, gci_rank)

```

```{r}
data <- gci_rankings_top_10 %>%
  # mutate(country = countrycode(country, origin = "country.name", destination = "iso2c")) %>%
    filter(year %in% c(2010:2018)) 

data %>% 
  ggplot(aes(year, gci_rank, group = country, colour = country, fill = country)) +
  geom_bump(aes(smooth = 20), size = 1.5, lineend = "round", alpha = 0.4) +
  geom_bump(data = data %>% filter(country == "United Kingdom"), 
            colour = "#d81159", size = 1.5, lineend = "round", linejoin = "bevel") +
#  geom_point(alpha = 0.4, size = 4) +
  scale_y_reverse(breaks = 1:10) +
# scale_colour_manual(values = mycolors) +
  scale_color_viridis_d(direction = -1, option = "plasma") +
  theme_minimal() +
  bump_theme() +
# add labels for 2010 rankings  
  geom_text(data = data %>% filter(year == 2010), 
            aes(label = paste(gci_rank, country), x = 0.5), 
            hjust = 1, colour = "grey40", fontface = "bold", size = 3) +
# add labels for 2018 rankings  
  geom_text(data = data %>% filter(year == 2018), 
            aes(label = paste(gci_rank, country), x = 9.5), 
            hjust = 0, colour = "grey40", fontface = "bold", size = 3) +
  scale_x_discrete(breaks = 2010:2018, expand = c(.28, .45)) +
  ggtitle("Country Rankings of Global Competitiveness Index") +
  ylim(c(12, 0))
```

The top 10 most competitive countries in 2018 have all been in the top 12 for the last decade, indicating that there is little change in rankings for these top performing countries. Sweden has fallen from 2nd to 9th. The United Kingdom has risen from 12th place to 9th. 


# Heatmap of top countries and each pillar ranking

```{r prepare data for heatmap}
# obtain the gci rankings since 2007 in order of gci
gci_top_10_countries <- gci_rankings %>%
  filter(year == 2018 & gci_rank <= 10) %>%
  arrange(gci_rank) %>%
  pull(country_code)

gci_top_10_country_names <- countrycode(gci_top_10_countries, origin = "iso3c", destination = "country.name")

# select out index values for 12 pillars
pillars <- gci_soc_env %>%
  # pillar values 
  select(country_code, starts_with("p_"), gci_overall) %>%
  # filter out countries of interest
  filter(country_code %in% gci_top_10_countries) %>%
  # add full country names
  mutate(country_name = countrycode(country_code, 
                                    origin = "iso3c", 
                                    destination = "country.name")) %>%
  # remove gci index total
  select(-gci_overall) %>%
  # pivot so that all pillar scores are in one column
  pivot_longer(cols = -c(country_name, country_code),
               names_to = "pillar",
               values_to = "score") %>%
  # modify pillar labels so that they are understandable for plot
  mutate(pillar_labs = (str_extract(pillar, "[a-z]*$")),
         pillar_labs = to_sentence_case(pillar_labs),
         pillar_labs = recode(pillar_labs,
           "Ict" = "ICT adoption",
           "Macroeconomic" = "Macroeconomic\nstability", 
           "Product" = "Product market", 
           "Labour" = "Labour market", 
           "Financial" = "Financial system", 
           "Market" = "Market size", 
           "Business" = "Business\ndynamism", 
           "Innovation" = "Innovation", 
           .default = pillar_labs
         ))


```



```{r gci_heatmap}
# function to replace the pillar names so that they are sequential for plot
for (x in 1:9){
  
  string <- paste("_", x, "_", sep = "")
  
  string_replace <- paste("_0", x, "_", sep = "")
  
  pillars <-  pillars %>%
  mutate(pillar = str_replace_all(pillar, string, string_replace))
  
}

# ORder the pillars in their order for factor reordering
pillar_order <- pillars %>%
  arrange(pillars) %>%
  distinct(pillar_labs) %>%
  pull(pillar_labs)

# Vector of colours of the 4 major categories of pillar
colours <- c(rep("#e04059", 4), rep("#eb701c", 2), rep("#4f529c", 4), rep("#38a380", 2))


# Heatmap of 12 pillars per 10 countries
pillars %>%
  #relevel pillars so they are plotted in order
  mutate(pillar_labs = fct_relevel(pillar_labs, pillar_order),
         country_name = fct_relevel(country_code, gci_top_10_countries)
        ) %>%
  ggplot() +
  geom_tile(aes(y = pillar_labs, 
                x = country_name, 
                fill = score),
            colour = "white", 
            size = 0.5) +
  # make the tiles square
  coord_equal() +
  # set the colour scale
  scale_fill_distiller(palette = "Purples",  
                       name = "Score", 
                       direction = 1,
                       guide = guide_colourbar(barwidth = 0.9, 
                                               barheight = 2.5,
                                               ticks = F,
                                               breaks = c(60, 100),
                                               labels = c(60, 100),
                                               label.position = "left",
                                               #title.position = "bottom",
                                               title.hjust = 1,
                                               title.vjust = 1.5)) +
  # add lines that group the pillars together
  # enabling environment
  annotate(
    geom = "line", y = c(0.6, 4.4), x = 0.3, col = colours[1], size = 2.5
  ) +
  # human capital
  annotate(
    geom = "line", y = c(4.6, 6.4), x = 0.3, col = colours[5], size = 2.5
  ) +
  # markets
  annotate(
    geom = "line", y = c(6.6, 10.4), x = 0.3, col = colours[8], size = 2.5
  ) +
  # innovation ecosystem
  annotate(
    geom = "line", y = c(10.6, 12.4), x = 0.3, col = colours[11], size = 2.5
  ) +
  # styling to remove all gridlines and axis
  theme_minimal() +
  # x axis label rotate 
  theme(axis.text.x = element_text(angle = 45, 
                                   #hjust = 1, # left justification
                                   #vjust = 0.5, 
                                   face = "bold"),
        
        axis.text.y = element_text(face = "bold",
                                   colour = colours),
        axis.title = element_text(size = 10,
                                  face = "bold"),
        axis.ticks = element_blank(),
        
        legend.title = element_text(size = 10, 
                                    face = "bold", 
                                    color = "grey40"),
        
        legend.text = element_text(face = "bold", 
                                   color = "grey40")) +
#  scale_x_discrete(position = "top") +
    labs(x = "Country Code", 
       y = NULL)

```


