---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(here)
library(broom)
library(factoextra)
```



The Global Competitiveness Index is an index that looks beyond short term measures and takes into account a full range of factors that determine long term productivity to assess a countries progress.

The index is made up of metrics classified under 12 pillars:
 * Institutions
 * Infrastructure
 * ICT adoption
 * Macroeconomic stability
 * Health
 * Skills
 * Product market
 * Labour market
 * Financial system
 * Market size
 * Business dynamism
 * Innovation capability.


"It is possible for an economy to be growing, inclusive and environmentally sustainable—but few economies are on such a trajectory"

"The GCI shows that other countries have very different results on social and environmental factors for the same level of current competitiveness and must begin work today to get on the path to creating not just growing but also low-carbon and inclusive economies."

I will begin by clutsering countries into broad groups based on the GCI. I will focus on the cluster containing the UK, and look at how their social and environmental policies compare to its counterparts. 

So you need to go back and include gci_overall, gdp_growth, measures of sustainability and social equity. First explore these factors. 

```{r}
ca_join <- read_csv(here("clean_data/gdp_gci_clustering.csv"))
```

# Deciding what variables to include for cluster analysis 
- economic growth
- GCI pillars


```{r}
# plot the different gdp metrics to decide whats important when comparing it to GCI
ca_join %>%
  pivot_longer(cols = starts_with("gdp"),
               names_to = "gdp_metric",
               values_to = "gdp_value") %>%
  ggplot() +
  aes(x = log(gdp_value), y = gci_overall) +
  geom_point() +
  facet_wrap(~gdp_metric, scales = "free_x")
```

I will select gdp per capital growth as a measure as it is not correlated to GCI, to investigate the groups. 

```{r}
# investigate the patterns within each pillar 
ca_join %>%
    pivot_longer(cols = starts_with("p_"),
               names_to = "pillar",
               values_to = "pillar_value") %>%
  ggplot() +
  aes(x = gdp_growth_annual, y = pillar_value, colour = country_code == "GBR") +
  geom_point(alpha = 0.8) +
  #geom_text(aes(label = country_code), check_overlap = TRUE) +
  facet_wrap(~pillar) +
  theme_bw() +
  scale_colour_viridis_d(option = "viridis", end = .8)+
  theme(legend.position = "none")
```

```{r}
library(broom)
library(corrr)

ca_join %>%
  select(-c(country_code, 
            starts_with("p_"), 
            ends_with("rank"), 
            gdp_per_capita_growth_annual, 
            gdp_growth_annual,
            total_consumption,
            gdp_ppp)) %>%
  correlate() %>%
  rearrange() %>%
  shave() %>%
  rplot(shape = 15) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```
Things that stand out from the correlation plot:

* countries with higher gdp have higher emissions
* stronger correlation of gdp to gci than gsci
* gdp correlated to happiness score
* gdp correlated slightly with gender gap index
* gdp is negatively correlated to gini - high income countries = poor distributino wealth
* some cor between renewable consumption as a proportion and ggi_index
* slight correlation between renewable energy growth and gdp annual growth
* growth per annum 08-18 is not correlated to growth in 19 - indicating there are countries which are either ramping up or ramping down renewables recently

```{r}
ca_join %>%
  ggplot +
  aes(x = ggi_index, y = renewable_cons_prop) +
  geom_point() +
  geom_text(aes(label = country_code))
```

```{r}
ca_join %>%
  ggplot +
  aes(x = rene_gen_growth_08_18, y = (ren_gen_growth_19)) +
  geom_point(aes(colour = country_code == "GBR")) +
  geom_text(aes(label = country_code), check_overlap = TRUE) +
  geom_abline(intercept = 0, slope = 1) +
  ylim(c(0, 1.1)) +
  theme(legend.position = "none")

```
Pakistan and Saudi reduced growth in renewables in 19
Venezuela and Vietnam are ramping up 


```{r}
ca_join %>%
  ggplot +
  aes(x = gci_overall, y = gsci_score) +
  geom_point() +
  geom_text(aes(label = country_code)) +
  geom_abline(intercept = 0, slope = 1, col = "red")

```

```{r}
ca_join %>%
    pivot_longer(cols = starts_with("gsci_"),
               names_to = "pillar",
               values_to = "pillar_value") %>%
  ggplot() +
  aes(x = gdp_per_capita_ppp, y = pillar_value, colour = country_code == "GBR") +
  geom_point(alpha = 0.8) +
  #geom_text(aes(label = country_code), check_overlap = TRUE) +
  facet_wrap(~pillar) +
  theme_bw() +
  scale_colour_viridis_d(option = "viridis", end = .8)+
  theme(legend.position = "none")
  
```





# Data preparation for clustering

```{r}
ca_join %>%
  is.na() %>%
  colSums()

ca_join %>%
  filter(is.na(gini_index))
```



```{r}
## Scale data and drop rows containing NAs
ca <- ca_join %>%
  mutate(across(is.numeric, ~as.numeric(scale(.))))


```


## Choosing K and validating clustering


```{r}
ca <- ca %>%
  drop_na(c(starts_with("gsci"), gdp_growth_decade)) %>%
  rename(overall_gsci = gsci_score)

ca_fviz <- select(ca, starts_with("gsci"))# , gdp_growth_decade) 

fviz_nbclust(ca_fviz, kmeans, method = "silhouette", nstart = 25)

```

```{r}
fviz_nbclust(ca_fviz, kmeans, method = "gap_stat") 
```

```{r}
fviz_nbclust(ca_fviz, kmeans, method = "wss", nstart = 25)

```

Select 3 as the optimal value of k

## Perform Kmeans clustering
```{r}
pillars_clust <- kmeans(ca_fviz, centers = 2)
```

```{r}
tidy(pillars_clust)
```

```{r}
# augment clusters onto data to explore other variables
ca_to_plot <- augment(pillars_clust, ca)


ca_to_plot %>%
  ggplot() +
  aes(y = overall_gsci, x = gdp_growth_decade, colour = .cluster) +
  geom_jitter(alpha = 0.6, width = 0.3) +
     scale_colour_viridis_d(option = "plasma", end = .9) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90)) +
  geom_text(aes(label = country_code), check_overlap = T)


```


```{r}
library(rworldmap)
library(rnaturalearth)
library(sf)
library(broom)

# obtain a table of countries and the clusters
ca_clusters <- ca_to_plot %>%
  select(country_code, .cluster) %>%
  distinct()

# obtain a sf object of the world
world <- ne_countries(scale = "small", returnclass = "sf")

# join clusters to world map
cluster_map <- world %>%
  full_join(ca_clusters, by = c("brk_a3" = "country_code"))

#plot clusters on world map
cluster_map %>%
ggplot() +
  geom_sf(aes(fill = cluster), size = 0.1, col = "black") +
  theme_void()+
  # change the projections mercator: 3395; robinson: 54030
  # find them all here: https://semba-blog.netlify.app/01/26/2020/world-map-and-map-projections/
  coord_sf(crs = 54030) +  
  scale_fill_manual(values = c("#E04059", "#4F529C"), name = "Cluster", breaks = c(1,2)) +
  theme(legend.position = c(0.2,0.4),
        legend.title = element_text(size = 8, face = "bold"),
          legend.key.size = unit(0.5, "cm"))


```





```{r}
ca_to_plot <- augment(pillars_clust, ca) %>%
    pivot_longer(cols = c(starts_with("gsci_"), gdp_per_capita_us),
               names_to = "pillar",
               values_to = "pillar_value")

  ggplot(ca_to_plot) +
  aes(x = overall_gsci, y = pillar_value, colour = .cluster) +
  geom_point() +
  facet_wrap(~pillar) +
  scale_colour_viridis_d(option = "viridis", end = .8)


```
Violin plot!

```{r}

# https://stackoverflow.com/questions/35717353/split-violin-plot-with-ggplot2

GeomSplitViolin <- ggproto(
  "GeomSplitViolin", 
  GeomViolin, 
  draw_group = function(self, data, ..., draw_quantiles = NULL) {
    data <- transform(data, 
                      xminv = x - violinwidth * (x - xmin), 
                      xmaxv = x + violinwidth * (xmax - x))
    grp <- data[1,'group']
    newdata <- plyr::arrange(
      transform(data, x = if(grp%%2==1) xminv else xmaxv), 
      if(grp%%2==1) y else -y
    )
    newdata <- rbind(newdata[1, ], newdata, newdata[nrow(newdata), ], newdata[1, ])
    newdata[c(1,nrow(newdata)-1,nrow(newdata)), 'x'] <- round(newdata[1, 'x']) 
    if (length(draw_quantiles) > 0 & !scales::zero_range(range(data$y))) {
      stopifnot(all(draw_quantiles >= 0), all(draw_quantiles <= 1))
      quantiles <- ggplot2:::create_quantile_segment_frame(data, draw_quantiles)
      aesthetics <- data[rep(1, nrow(quantiles)), setdiff(names(data), c("x", "y")), drop = FALSE]
      aesthetics$alpha <- rep(1, nrow(quantiles))
      both <- cbind(quantiles, aesthetics)
      quantile_grob <- GeomPath$draw_panel(both, ...)
      ggplot2:::ggname("geom_split_violin", 
                       grid::grobTree(GeomPolygon$draw_panel(newdata, ...), quantile_grob))
    } else {
      ggplot2:::ggname("geom_split_violin", GeomPolygon$draw_panel(newdata, ...))
    }
  }
)

geom_split_violin <- function (mapping = NULL, 
                               data = NULL, 
                               stat = "ydensity", 
                               position = "identity", ..., 
                               draw_quantiles = NULL, 
                               trim = TRUE, 
                               scale = "area", 
                               na.rm = FALSE, 
                               show.legend = NA, 
                               inherit.aes = TRUE) {
  layer(data = data, 
        mapping = mapping, 
        stat = stat, 
        geom = GeomSplitViolin, 
        position = position, 
        show.legend = show.legend, 
        inherit.aes = inherit.aes, 
        params = list(trim = trim, 
                      scale = scale, 
                      draw_quantiles = draw_quantiles, 
                      na.rm = na.rm, ...)
        )
}
```


```{r}

ca_join %>%
  left_join(ca_clusters, by = "country_code") %>%
  pivot_longer(cols = c(starts_with("gsci_")),
               names_to = "pillar",
               values_to = "pillar_value") %>%
  arrange(.cluster) %>%
  ggplot() +
  aes(x = pillar, y = pillar_value, fill = .cluster) +
  geom_split_violin(alpha = 0.9) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = c("#E04059", "#38A380"))
```

```{r}
ca_join %>%
  left_join(ca_clusters, by = "country_code") %>%
  ggplot() +
  aes(x = .cluster, y = gdp_per_capita_ppp, fill = .cluster) +
  geom_violin(alpha = 0.9) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_manual(values = c("#E04059", "#38A380"))
```





The two groups of countries are as such:

Cluster 1 low governance and intellectual capital and social capital scores.

Cluster 2 higher governance, intellectual capital and social capital. 

Natural capital and resource intensity showed little difference between the two




Results of the clustering:

Two groups are

```{r}
ca_to_plot <- augment(pillars_clust, ca) %>%
  pivot_longer(cols = starts_with("gdp"),
               names_to = "gdp_metric",
               values_to = "gdp_value")
 
  
   ggplot(ca_to_plot) +
  aes(x = (gdp_value), y = gci_overall, colour = .cluster) +
  geom_point() +
    geom_point(data = subset(ca_to_plot, country_code == "GBR"), aes(x = (gdp_value), y =  gci_overall), colour = "black") +
  facet_wrap(~gdp_metric, scales = "free_x") +
     scale_colour_viridis_d(option = "viridis", end = .8)
```

Analysis indicates:

Countries seem to be stratified accross all pillars, though there are some anomalies. The data is perhaps not best suited to clustering, however the groupings can be used for comparisons.

The scores within each pillar appear to be correlated, and there appears to be correlation between 

In terms of growth 



```{r}
ca_join_plot <- ca_join %>%
  pivot_longer(cols = starts_with("p_"),
               names_to = "pillar",
               values_to = "pillar_value")  

ca_join_plot %>%
  ggplot() +
  aes(x = pillar, y = pillar_value, colour = cut(gdp_growth_annual, c(-Inf, 0, 3, 5, Inf))) +
  geom_jitter(alpha = 0.6, width = 0.3) +
     scale_colour_viridis_d(option = "plasma", end = .9) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "bottom") +
  geom_point(data = subset(ca_join_plot, country_code == "GBR"), 
             aes(x = pillar, 
                 y =  pillar_value), colour = "black")



```


