---
title: "GSCI Clustering"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```


```{r}
library(tidyverse)
library(here)
library(broom)
library(factoextra)
library(rworldmap)
library(rnaturalearth)
library(sf)
library(corrr)
library(ggridges)
```


```{r}
ca_join <- read_csv(here("clean_data/gdp_gci_clustering.csv"))
```


# Correlation plot

```{r}
ca_join %>%
  select(-c(country_code, 
            starts_with("p_"), 
            ends_with("rank"), 
            gdp_per_capita_growth_annual, 
            gdp_growth_annual,
            total_consumption,
            gdp_ppp)) %>%
  correlate() %>%
  rearrange() %>%
  shave() %>%
  rplot(shape = 15) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```
Things that stand out from the correlation plot:

* countries with higher gdp have higher emissions
* stronger correlation of gdp to gci than gsci
* gdp correlated to happiness score
* gdp correlated slightly with gender gap index
* gdp is negatively correlated to gini - high income countries = poor distributino wealth
* some cor between renewable consumption as a proportion and ggi_index
* slight correlation between renewable energy growth and gdp annual growth
* growth per annum 08-18 is not correlated to growth in 19 - indicating there are countries which are either ramping up or ramping down renewables recently



```{r}
ca_join %>%
    pivot_longer(cols = starts_with("gsci_"),
               names_to = "pillar",
               values_to = "pillar_value") %>%
  ggplot() +
  aes(x = gdp_per_capita_ppp, y = pillar_value, colour = country_code == "GBR") +
  geom_point(alpha = 0.8) +
  #geom_text(aes(label = country_code), check_overlap = TRUE) +
  facet_wrap(~pillar) +
  theme_bw() +
  scale_colour_viridis_d(option = "viridis", end = .8)+
  theme(legend.position = "none")
  
```

Decided not to include gdp into clustering, but will just use the gsci index 


# Cluster Analysis 

## Data preparation for clustering

```{r}
ca_join %>%
  is.na() %>%
  colSums()
```



```{r}
## Scale data 
ca <- ca_join %>%
  mutate(across(where(is.numeric), ~as.numeric(scale(.))))

```


## Choosing K and validating clustering


```{r}
ca <- ca %>%
  drop_na(c(starts_with("gsci"), gdp_growth_decade)) %>%
  rename(overall_gsci = gsci_score)

ca_fviz <- select(ca, starts_with("gsci"))# , gdp_growth_decade) 

fviz_nbclust(ca_fviz, kmeans, method = "silhouette", nstart = 25)

```

```{r}
fviz_nbclust(ca_fviz, kmeans, method = "gap_stat") 
```

```{r}
fviz_nbclust(ca_fviz, kmeans, method = "wss", nstart = 25)

```

Select 2 as the optimal number of cluster


## Perform Kmeans clustering
```{r}
pillars_clust <- kmeans(ca_fviz, centers = 2)
```


### Model output

```{r}
tidy(pillars_clust)
```

```{r}
# augment clusters onto data to explore other variables
ca_to_plot <- augment(pillars_clust, ca)
```


```{r}

# obtain a table of countries and the clusters
ca_clusters <-
  ca_to_plot %>%
  select(country_code, .cluster) %>%
  distinct()

write_csv(ca_clusters, here("clean_data/clusters.csv"))

# obtain a sf object of the world
world <- ne_countries(scale = "small", returnclass = "sf")

# join clusters to world map
cluster_map <- world %>%
  full_join(ca_clusters, by = c("brk_a3" = "country_code"))

#plot clusters on world map
#map <- 
cluster_map %>%
ggplot() +
  geom_sf(aes(fill = .cluster), size = 0.1, col = "black") +
  theme_void()+
  # change the projections mercator: 3395; robinson: 54030
  # find them all here: https://semba-blog.netlify.app/01/26/2020/world-map-and-map-projections/
  coord_sf(crs = 54030) +  
  scale_fill_manual(values = c("#E04059", "#4F529C"), name = "Cluster", breaks = c(1,2)) +
  theme(legend.position = c(0.2,0.4),
        legend.title = element_text(size = 8, face = "bold"),
          legend.key.size = unit(0.5, "cm")) +
#  labs(title = "Countries Grouped According to Global Sustainable Competitiveness") +
  theme(plot.title = element_text(color="grey40", size=15, face = "bold"))

#ggsave(here("plot_images/cluster_map.png"), map)


```

# Pillar Scores for each cluster

```{r}
ridge_plot <-  ca_join %>%
  left_join(ca_clusters, by = "country_code") %>%
  select(-gsci_score) %>% 
    pivot_longer(cols = c(starts_with("gsci_")),
               names_to = "pillar",
               values_to = "pillar_value") %>%
  ggplot() +
  aes(x = pillar_value, 
      y = pillar, 
      color = .cluster, 
      fill = .cluster) +
    stat_density_ridges(quantile_lines = TRUE, 
                        quantiles = 2, 
                        alpha = 0.6, 
                        scale = 0.95, 
                        rel_min_height = .01) +
    # scale_fill_manual(values = c("#E04059", "#4F529C"), name = "Cluster") +
    # scale_color_manual(values = c("#E04059", "#4F529C"), name = "Cluster") +
    theme_ridges(center_axis_labels = TRUE) +
  labs(x = "Score") +
  scale_y_discrete(labels = c("Governance", "Intellectual\nCapital", 
                              "Natural Capital", "Resource\nIntensity", "Social Capital")) +
  scale_x_continuous(breaks = c(seq(10, 80, by = 10))) +
    theme(
    axis.title.y = element_blank(),
    axis.text.x = element_text(face = "bold"),
    plot.title = element_text(face = "bold"),
    axis.text.y = element_text(face = "bold", hjust = 1),
    plot.margin = unit(c(1.5, 1, 1.5, 1), "cm") #top, right, bottom, left
  )

# ggsave(here("plot_images/cluster_ridge.png"), ridge_plot, height = 6)
```

The two groups of countries are as such:

Cluster 1 low governance, intellectual capital and social capital scores.

Cluster 2 higher governance, intellectual capital and social capital. 

Natural capital and resource intensity showed little difference between the two




```{r}
ca_to_plot <- augment(pillars_clust, ca) %>%
  pivot_longer(cols = starts_with("gdp"),
               names_to = "gdp_metric",
               values_to = "gdp_value")
 
  
   ggplot(ca_to_plot) +
  aes(x = (gdp_value), y = gci_overall, colour = .cluster) +
  geom_point() +
    geom_point(data = subset(ca_to_plot, country_code == "GBR"), aes(x = (gdp_value), y =  gci_overall), colour = "black") +
  facet_wrap(~gdp_metric, scales = "free_x") +
     scale_colour_viridis_d(option = "viridis", end = .8)
```




